RDA trait prediction tutorial

### 2025-10-16 RDAtraitPredictionTutorial.Rmd
modified RDA tutorial function to be compatible with newer version of vegan:

```{r function}
rda_trait_pred <- function(rdaobj, env_row, K){
  #rdaobj is RDA object
  #env_row is the row of the environment in the biplot output
  #K is the number of RDA axes
  scores <- scores(rdaobj, choices=1:K)
  ind.sc <- scores$sites
  pred <- matrix(NA, nrow=nrow(ind.sc), ncol=K)
  for (k in 1:K){
    pred[,k] <- ind.sc[,k]*eigenvals(rdaobj)[k]*summary(rdaobj)$biplot[env_row,k]
  }
  trait_pred <- scale(rowSums(pred))
 return(trait_pred) 
}
```

>>>>

```{r function}
#### old function from Katie with older version of vegan
# rda_trait_pred <- function(rdaobj, env_row, K){
#   #rdaobj is RDA object
#   #env_row is the row of the environment in the biplot output
#   #K is the number of RDA axes
#   scores <- scores(rdaobj, choices=1:K)
#   ind.sc <- scores$sites
#   pred <- matrix(NA, nrow=nrow(ind.sc), ncol=K)
#   for (k in 1:K){
#     pred[,k] <- ind.sc[,k]*eigenvals(rdaobj)[k]*summary(rdaobj)$biplot[env_row,k]
#   }
#   trait_pred <- scale(rowSums(pred))
#  return(trait_pred) 
# }

##### updated function that works with newer version of vegan
rda_trait_pred <- function(rdaobj, env_row, K){
  # clamp K to available constrained axes
  K <- min(K, length(vegan::eigenvals(rdaobj)))

  # site scores and biplot loadings for the first K axes
  ind.sc <- vegan::scores(rdaobj, display = "sites", choices = 1:K)
  bp     <- vegan::scores(rdaobj, display = "bp",    choices = 1:K)

  pred <- matrix(NA, nrow = nrow(ind.sc), ncol = K)
  for (k in 1:K){
    pred[, k] <- ind.sc[, k] * vegan::eigenvals(rdaobj)[k] * bp[env_row, k]
  }
  trait_pred <- scale(rowSums(pred))
  return(trait_pred)
}
```
